Supplementary Information

### 2015-06-24 manuscript/latitudepaper.tex
The Brose webs aren't Brose webs

  %Food webs are a thing. (Mention gen, vul, LS.) %Food webs have strong scaling structure.
  Food webs --networks of feeding links between species-- have been used for several decades to summarise the structure of 
  ecological communities~\citep{Petchey2008,Williams2000,Paine1966} and to understand how that structure relates to environmental variables 
  such as habitat type~\citep{Shurin2006,Briand1983}, primary productivity~\citep{Vermaat2009,Thompson2005a,Townsend1998}, and climate~\citep{Baiser2012,Petchey2010}. The latter
  variables in turn have strong gradients over latitude, with productivity and temperature both being higher in the tropics while 
  climate is more variable at high latitudes~\citep{Field1998}. These variables affect both the resources available and species' 
  metabolisms~\citep{Hechinger2011,White2011,OConnor2009,White2007}, and have been proposed as determinants of the strength of 
  interspecific interactions~\citep{Lang2012,Schleuning2012,Schemske2009}. 
  By modulating interactions between species, latitudinal gradients may also shape food-web structure.
  Indeed, these latitudinal environmental gradients have been put forward as potential drivers for the 
  latitudinal gradient in species richness, one of the most general and robust patterns in ecology~\citep{Schemske2009,Macpherson2002,Kaufman1995}.

@@

  If the latitude-niche breadth hypothesis is correct, there should also be direct relationships between latitude and the
  degree of specialisation (i.e., the breadth of the Eltonian niche ~\citep{Elton1927}) of species within food webs.
  Attempts to unravel these effects, however,
  are complicated by known relationships between species richness and many other network properties~\citep{Riede2010}. 
  For example,
  narrower niches imply fewer links per species (i.e., greater specialisation) in the tropics(~\citealp{Dyer2007,Marra1997};
  but see~\citealp{Schleuning2012}). However,
  average numbers of links per species tend to increase in larger food webs~\citep{Dunne2006,Riede2010}. This means
  that a latitudinal effect on specialisation may be obscured by a latitudinal gradient in species richness. If this is
  the case, it may still be possible to uncover effects of latitude on specialisation by examining the shape of the scaling 
  relationship between specialisation and species richness over changing latitude. Here, we use three measures of specialisation:
  mean links per species, mean generality (number of prey), and mean vulnerability (number of predators). By testing whether
  latitude affects the scaling of each property with species richness, we test effects of latitude on specialisation implied by
  the ``latitude-niche breadth hypothesis''.

@@

    As the food webs in this dataset are derived from a variety of sources and were compiled over many decades, it
    is likely that they vary in their resolution and in the amount of sampling effort invested in their assembly.
    Many analyses of food-web structure attempt to reduce this variation by aggregating species with identical predator and prey
    sets to form ``trophic species'' webs ~\citep{Martinez1991,Vermaat2009,Dunne2004,Dunne2013}. As our study
    is concerned directly with the number of species at a particular latitude, however, we did not wish to ignore 
    species with identical sets of interactions. We therefore analysed both original and trophic-species versions
    of the dataset; in each case using the number of species and 
    links in each web to calculate the mean link density (number of links per species), mean generality 
    (number of prey per species), and mean vulnerability (number of predators per species) of the web. 
    The version of the dataset used did not qualitatively change the results, suggesting that
    the scaling relationships between species richness, other food-web properties, and latitude are very 
    similar whether or not redundant species are included. For simplicity, here we present only
    the results for the original webs.

@@

  The tendency of food-web structure to exhibit scaling relationships with
  species richness has been well-established~\citep{Dunne2004,Riede2010}. As
  species richness in particular is also known to vary systematically over
  latitude~\citep{Schemske2009,Macpherson2002,Kaufman1995}, intuitively one
  might suspect that any relationship between food-web properties such as
  generality might be due to the latitudinal gradient in species richness. In
  this dataset, however, we did not find overall latitudinal gradients in
  species richness, links per species,  generality, vulnerability, or the
  proportions of food webs accounted for by  basal resources, intermediate
  consumers, and top predators (see \emph{Appendix S3} for details).

  The lack of a latitudinal gradient in species richness in this dataset
  contrasts strongly with other
  studies~\citep{Schemske2009,Macpherson2002,Kaufman1995}. As numbers of
  species and links included in a food web vary strongly with sampling effort
  as well as with the underlying diversity of the study area, it is possible
  that the lack of latitudinal trends here is a result of researchers tending
  to expend similar amounts of sampling effort across studies. This could
  result in food webs describing species-rich tropical communities omitting
  more species and links than species-poor arctic communities. In addition, it
  is worth noting that gradients in species richness are generally measured
  for a single taxonomic group at a
  time~\citep[e.g.]{Schemske2009,Macpherson2002,Kaufman1995}. It is possible
  that these taxa are not well-represented in our food webs and that the
  dominant taxa in them do not have an underlying latitudinal gradient in
  richness.In either case, the lack of a strong association between species
  richness and latitude in any ecosystem type means that any effect of
  latitude on other scalingrelationships is not being driven by an
  underlyinglatitudinal gradient in species richness. The lack of confounding
  effects of latitude allows us tomore clearly assess effects of latitude on
  scaling with species richness and proportions of species in differenttrophic
  levels.

@@

  Unlike other ecosystem types, the scaling of generality in lake and stream
  food webs was stronger (i.e.,generality increased more steeply with
  increasing species richness) in higher-latitude food webs. In lake food webs,
  this trend was echoed in the scaling relationships between species richness
  and links per species and vulnerability. This means that species in tropical
  freshwater communities gain fewer additional feeding links per additional
  species in the web and that species in tropical lakes also gain fewer
  predators, and fewer links in general, per additional species than species
  in high-latitude lakes. These trends are consistent with the hypothesis that
  greater stability in the tropics leads to narrower
  niches~\citep{Vazquez2004} and a higher proportion of specialists. This may
  be partly due to high-latitude species tending to switch between different
  seasonally-available prey~\citep{Isaac2012,Wilhelm1999} while tropical
  freshwater ecosystems may have more stable composition.

  The scaling of link density, generality and vulnerability with proportions
  of species at different trophic levels, however, did not differ between lakes
  and most other food webs. Indeed, there was comparatively little variation in
  scaling with the trophic-level breakdown of a web across ecosystem types.
  Terrestrial food webs tended to exhibit less change in scaling over latitude
  than other ecosystem types, as did marine webs when considering scaling with
  the proportion of basal resources. More excitingly, the negative relationships
  between the proportion of top predators in a web and link density and
  vulnerability were weaker (i.e., the scaling exponent was more strongly
  negative) in low-latitude stream ecosystems. This may imply that top predators
  in high-latitude streams tend to be more generalist than their low-latitude
  equivalents~\citep{Winemiller2008}.
  More generally, the lack of correspondence between the scaling of food-web
  properties with species richness and with proportions of species at each
  trophic level suggests that the size and trophic breakdowns of a community can
  provide different information about the breadth of niches in that community.

>>>>

  %Food webs are a thing. (Mention gen, vul, LS.) %Food webs have strong scaling structure.

  Food webs --networks of feeding links between species-- have been used for
  several decades to summarise the structure of  ecological
  communities~\citep{Paine1966,Williams2000,Petchey2008} and to understand how
  that structure relates to environmental variables  such as habitat
  type~\citep{Briand1983,Shurin2006}, primary
  productivity~\citep{Townsend1998,Thompson2005a,Vermaat2009}, and
  climate~\citep{Petchey2010,Baiser2012}. The latter variables in turn have
  strong gradients over latitude, with productivity and temperature both being
  higher in the tropics while  climate is more variable at high
  latitudes~\citep{Field1998}. These variables affect both the resources
  available and species'
  metabolisms~\citep{White2007,OConnor2009,Hechinger2011,White2011}, and have
  been proposed as determinants of the strength of  interspecific
  interactions~\citep{Schemske2009,Lang2012,Schleuning2012}.  By modulating
  interactions between species, latitudinal gradients may also shape food-web
  structure. Indeed, these latitudinal environmental gradients have been put
  forward as potential drivers for the  latitudinal gradient in species
  richness, one of the most general and robust patterns in
  ecology~\citep{Kaufman1995,Macpherson2002,Schemske2009}.

@@

  If the latitude-niche breadth hypothesis is correct, there should also be direct relationships between latitude and the
  degree of specialisation (i.e., the breadth of the Eltonian niche ~\citep{Elton1927}) of species within food webs.
  Attempts to unravel these effects, however,
  are complicated by known relationships between species richness and many other network properties~\citep{Riede2010}. 
  For example,
  narrower niches imply fewer links per species (i.e., greater specialisation) in the tropics(~\citealp{Marra1997,Dyer2007};
  but see~\citealp{Schleuning2012}). However,
  average numbers of links per species tend to increase in larger food webs~\citep{Dunne2006,Riede2010}. This means
  that a latitudinal effect on specialisation may be obscured by a latitudinal gradient in species richness. If this is
  the case, it may still be possible to uncover effects of latitude on specialisation by examining the shape of the scaling 
  relationship between specialisation and species richness over changing latitude. Here, we use three measures of specialisation:
  mean links per species, mean generality (number of prey), and mean vulnerability (number of predators). By testing whether
  latitude affects the scaling of each property with species richness, we test effects of latitude on specialisation implied by
  the ``latitude-niche breadth hypothesis''.

@@

    As the food webs in this dataset are derived from a variety of sources and were compiled over many decades, it
    is likely that they vary in their resolution and in the amount of sampling effort invested in their assembly.
    Many analyses of food-web structure attempt to reduce this variation by aggregating species with identical predator and prey
    sets to form ``trophic species'' webs ~\citep{Martinez1991,Dunne2004,Vermaat2009,Dunne2013}. As our study
    is concerned directly with the number of species at a particular latitude, however, we did not wish to ignore 
    species with identical sets of interactions. We therefore analysed both original and trophic-species versions
    of the dataset; in each case using the number of species and 
    links in each web to calculate the mean link density (number of links per species), mean generality 
    (number of prey per species), and mean vulnerability (number of predators per species) of the web. 
    The version of the dataset used did not qualitatively change the results, suggesting that
    the scaling relationships between species richness, other food-web properties, and latitude are very 
    similar whether or not redundant species are included. For simplicity, here we present only
    the results for the original webs.

@@

  The tendency of food-web structure to exhibit scaling relationships with
  species richness has been well-established~\citep{Dunne2004,Riede2010}. As
  species richness in particular is also known to vary systematically over
  latitude~\citep{Kaufman1995,Schemske2009,Macpherson2002}, intuitively one
  might suspect that any relationship between food-web properties such as
  generality might be due to the latitudinal gradient in species richness. In
  this dataset, however, we did not find overall latitudinal gradients in
  species richness, links per species,  generality, vulnerability, or the
  proportions of food webs accounted for by  basal resources, intermediate
  consumers, and top predators (see \emph{Appendix S3} for details).

  The lack of a latitudinal gradient in species richness in this dataset
  contrasts strongly with other
  studies~\citep{Kaufman1995,Schemske2009,Macpherson2002}. As numbers of
  species and links included in a food web vary strongly with sampling effort
  as well as with the underlying diversity of the study area, it is possible
  that the lack of latitudinal trends here is a result of researchers tending
  to expend similar amounts of sampling effort across studies. This could
  result in food webs describing species-rich tropical communities omitting
  more species and links than species-poor arctic communities. In addition, it
  is worth noting that gradients in species richness are generally measured
  for a single taxonomic group at a
  time~\citep[e.g.]{Kaufman1995,Schemske2009,Macpherson2002}. It is possible
  that these taxa are not well-represented in our food webs and that the
  dominant taxa in them do not have an underlying latitudinal gradient in
  richness.In either case, the lack of a strong association between species
  richness and latitude in any ecosystem type means that any effect of
  latitude on other scalingrelationships is not being driven by an
  underlyinglatitudinal gradient in species richness. The lack of confounding
  effects of latitude allows us tomore clearly assess effects of latitude on
  scaling with species richness and proportions of species in differenttrophic
  levels.

@@

  Unlike other ecosystem types, the scaling of generality in lake and stream
  food webs was stronger (i.e.,generality increased more steeply with
  increasing species richness) in higher-latitude food webs. In lake food webs,
  this trend was echoed in the scaling relationships between species richness
  and links per species and vulnerability. This means that species in tropical
  freshwater communities gain fewer additional feeding links per additional
  species in the web and that species in tropical lakes also gain fewer
  predators, and fewer links in general, per additional species than species
  in high-latitude lakes. These trends are consistent with the hypothesis that
  greater stability in the tropics leads to narrower
  niches~\citep{Vazquez2004} and a higher proportion of specialists. This may
  be partly due to high-latitude species tending to switch between different
  seasonally-available prey~\citep{Magalhaes1993,Wilhelm1999,Isaac2012} while tropical
  freshwater ecosystems may have more stable composition.

  The scaling of link density, generality and vulnerability with proportions
  of species at different trophic levels, however, did not differ between lakes
  and most other food webs. Indeed, there was comparatively little variation in
  scaling with the trophic-level breakdown of a web across ecosystem types.
  Terrestrial food webs tended to exhibit less change in scaling over latitude
  than other ecosystem types, as did marine webs when considering scaling with
  the proportion of basal resources. More excitingly, the negative relationships
  between the proportion of top predators in a web and link density and
  vulnerability were weaker (i.e., the scaling exponent was more strongly
  negative) in low-latitude stream ecosystems. This may imply that top predators
  in high-latitude streams tend to be more generalist than their low-latitude
  equivalents~\citep{Winemiller2008}, perhaps due to prey-switching during seasonal
  food shortages~\citep{Magalhaes1993}.
  More generally, the lack of correspondence between the scaling of food-web
  properties with species richness and with proportions of species at each
  trophic level suggests that the size and trophic breakdowns of a community can
  each provide different information~\citep{Downing2002}.

### 2015-06-24 manuscript/supplement.tex
The Brose webs aren't Brose webs

  We combined published food webs from~\citet{GlobaWeb},~\citet{Brose2006}, 
  and~\citet{Dunne2013}.
  Of the 385 food webs available from~\citet{Globalweb}, 122 were rejected because their original
  source was unpublished or could not be retrieved. We further eliminated 13 source or sink webs
  (which describe feeding links flowing from or to a particular resource or top predator rather
  than the interactions of an entire community), 8 plant-pollinator webs, 34 host-parasitoid webs,
  and 2 networks describing competitive interactions. Finally, we eliminated 2 webs describing 
  inferred interactions in an extinct community, 8 ``generalized schemes'' that were not based
  on direct field data, and 2 webs where it was not clear which of several described study sites
  were used to construct the food web (and therefore the latitude of the food web could not be 
  included). This left us with 156 webs from~\citet{GlobalWeb}.

  To these, we added 7 webs from~\citet{Dunne2013} that were originally used to assess the roles
  of parasites within food webs. As all but 3 of the webs from~\citet{GlobalWeb} contained only
  free-living species, we removed the the parasites (and all interactions involving them) from
  these webs (and the 3 webs containing parasites from~\citet{GlobalWeb}) leaving free-living 
  species only. In addition, we included 12 of the 96 webs from~\citet{Brose2005}, which were originally
  collected to test the relationship between consumer and
  resources body sizes. The remaining webs the~\citet{Brose2005}
  dataset were also included in the~\citet{GlobalWeb} 
  dataset. This gave us a database of 259 webs. As we
  wished to use the logarithmic form of power-law
  relationships, we then eliminated any food web which
  did not have at least one each of basal resources,
  intermediate consumers, and top predators (to obtain 
  non-zero proportions of each category). Our final sample
  size was thereby reduced to 163 food webs.

>>>>

  We combined published food webs from~\citet{GlobaWeb},~\citet{Riede2011}, 
  and~\citet{Dunne2013}.
  Of the 385 food webs available from~\citet{Globalweb}, 122 were rejected because their original
  source was unpublished or could not be retrieved. We further eliminated 13 source or sink webs
  (which describe feeding links flowing from or to a particular resource or top predator rather
  than the interactions of an entire community), 8 plant-pollinator webs, 34 host-parasitoid webs,
  and 2 networks describing competitive interactions. Finally, we eliminated 2 webs describing 
  inferred interactions in an extinct community, 8 ``generalized schemes'' that were not based
  on direct field data, and 2 webs where it was not clear which of several described study sites
  were used to construct the food web (and therefore the latitude of the food web could not be 
  included). This left us with 156 webs from~\citet{GlobalWeb}.

  To these, we added 7 webs from~\citet{Dunne2013} that were originally used to assess the roles
  of parasites within food webs. As all but 3 of the webs from~\citet{GlobalWeb} contained only
  free-living species, we removed the the parasites (and all interactions involving them) from
  these webs (and the 3 webs containing parasites from~\citet{GlobalWeb}) leaving free-living 
  species only. In addition, we included 30 of the 35 webs from~\citet{Riede2011}, which were originally
  collected to test the relationship between consumer and
  resources body sizes. The remaining webs the~\citet{Brose2005}
  dataset were also included in the~\citet{GlobalWeb} 
  dataset. This gave us a database of 264 webs. As we
  wished to use the logarithmic form of power-law
  relationships, we then eliminated any food web which
  did not have at least one each of basal resources,
  intermediate consumers, and top predators (to obtain 
  non-zero proportions of each category). Our final sample
  size was thereby reduced to 197 food webs.

% Tree hole and pitcher plant webs counted as "lakes", wetlands could be either lakesor streams depending on the study
